---
title: "Statistical assignment 1"
author: "Simone Long_135288"
date: "03/02/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open data (10 points)

```{r message=FALSE, warning=FALSE}
library(tidyverse)
Data <- read_tsv("data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
```

## Select variables (10 points)

```{r}
Data <- Data %>%
        select(pidp, h_eumem, h_sex_dv, h_age_dv, h_memorig)
```

## Filter observations (10 points)

```{r}
Data <- Data %>%
        filter(h_memorig == 1)
```

## Recode data (20 points)

```{r}
table(Data$h_eumem)
table(Data$h_sex_dv)
table(Data$h_age_dv)
```

```{r message=FALSE, warning=FALSE}
Data <- Data %>%
        mutate(EU = ifelse(h_eumem == 1, 1,
                           ifelse(h_eumem == 2, 0, NA))) %>%
        mutate(sex = recode(h_sex_dv,
                `1` = "male",
                `2` = "female")) %>%
        mutate(agegr = case_when(
          between(h_age_dv, 16, 25) ~ "16 to 25",
          between(h_age_dv, 26, 40) ~ "26 to 40",
          between(h_age_dv, 41, 55) ~ "41 to 55",
          between(h_age_dv, 56, 70) ~ "56 to 70",
          h_age_dv > 70 ~ "over 70"
        ))
```

```{r}
EUSupport2 <- table(Data$EU)
knitr::kable(EUSupport2,
             caption = "*EU Support: Recoded*")
Sex2 <- table(Data$sex)
knitr::kable(Sex2,
             caption = "*Sex: Recoded*")
Age2 <- table(Data$agegr)
knitr::kable(Age2,
             caption = "*Age: Recoded*")
```


## Summarise data (20 points)

```{r}
Data %>%
        count(EU) %>%
        mutate(perc = n / sum(n) * 100)
        
```

Based on this sample, the majority of UK voters should have voted to Remain in the UK in the 2016 Referendum, which, of course, did not happen. One reason for this error in prediction could be the percentage of missing responses (which, in this sample, amounts to nearly 11%). It is also likely that, because the data collection spanned 3 years, the opinions of some sample members changed since their initial vote in the Referendum. But the sampling technique used in this study could be to blame; participation is completely voluntary, time-consuming, and necessitates the ownership of a residence/phone/internet/etc for completion. Therefore, it is probable that the sample is not entirely representative of the UK, which could explain the difference between the study's findings and the actual result of the Referendum.

## Summarise data by sex and age (30 points)

```{r}
Data %>%
        group_by(sex, agegr) %>%
        summarise(
          propLeave = mean(EU == 0, na.rm = TRUE) *  100, 
          propRemain = mean(EU == 1, na.rm = TRUE) * 100, 
          totalLeave = sum(EU == 0, na.rm = TRUE),
          totalRemain = sum(EU == 1, na.rm = TRUE)
        )
        
```

Based on this sample in Great Britain, it appears that women and men aged 16-25 were most likely to have voted to remain in the EU, with women generally expressing a greater likelihood of voting to remain in all age groups. However, age and voting to leave the EU seem to be positively correlated, meaning that those aged over 70 tended to choose to leave, with older men showing consistently greater interest in leaving than older women.